---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

Let's load the data and see the head of it.

```{r,echo=TRUE,comment=''}
Activity<-read.csv('./activity/activity.csv')
Activity$date<-as.Date(Activity$date)
head(Activity)
```


## What is mean total number of steps taken per day?

First, we are going to calculate the total number of steps per day.

```{r,echo=TRUE,comment=''}
Steps<-aggregate(Activity$steps,list(Activity$date),sum,na.rm=TRUE)
colnames(Steps)<-c('Day','Steps')
Steps
```

Let's plot the histogram.

```{r,echo=TRUE,comment=''}
hist(Steps$Steps,xlab='Number of steps taken every day',main='Steps')
```

Now let's calculate the mean and the median of steps per day

```{r,echo=TRUE,comment=''}
MeanMedian<-aggregate(x=Activity$steps,by=list(Activity$date),FUN=function(x) c(media=mean(x,na.rm = TRUE), mediana=median(x,na.rm=TRUE)))
MeanMedian
```


## What is the average daily activity pattern?

Let's calculate the Average daily activity as a time series.

```{r,echo=TRUE,comment=''}
Pattern<-aggregate(Activity$steps,list(Activity$interval),mean,na.rm=TRUE)
colnames(Pattern)<-c('Interval','Steps')
plot(Pattern$Interval,Pattern$Steps,type = 'l',xlab='Intervals',ylab='Steps',main='Average daily activity')
```

The 5-minute interval which contains the maximum number of steps n average across all the days is:

```{r,echo=TRUE,comment=''}
Pattern[which.max(Pattern$Steps),'Interval']
```


## Imputing missing values

Let's calculate the total of rows with Nas

```{r,echo=TRUE}
sum(rowSums(is.na(Activity))>0)
```


Let's fill the NA's of the steps with the mean of the intervals and create a new dataset

```{r,echo=TRUE}
NewActivity<-Activity
dates<-unique(NewActivity$date)

for(i in dates){
  if(sum(is.na(NewActivity[which(NewActivity$date==i),]$steps))==nrow(NewActivity[which(NewActivity$date==i),])){
    NewActivity[which(NewActivity$date==i),]$steps<-Pattern$Steps
  }
}
NewActivity$date<-as.Date(NewActivity$date)
```

Let's see the new total of steps per day and plot the new histogram.

```{r,echo=TRUE}
NewSteps<-aggregate(NewActivity$steps,list(NewActivity$date),sum,na.rm=TRUE)
colnames(NewSteps)<-c('Day','Steps')
NewSteps
```

```{r,echo=TRUE}
hist(NewSteps$Steps,xlab='Number of steps taken every day',main='Steps')
```

Now, let's calculate the mean and median of the new dataset.

```{r,echo=TRUE,comment=''}
NewMeanMedian<-aggregate(x=NewActivity$steps,by=list(NewActivity$date),FUN=function(x) c(media=mean(x,na.rm = TRUE), mediana=median(x,na.rm=TRUE)))
NewMeanMedian
```

We can tell that the steps between 10,000 and 15,000 had a increased

## Are there differences in activity patterns between weekdays and weekends?

Let's create a new factor variable if it is weekday or weekend

```{r,echo=TRUE}
days<-weekdays(NewActivity$date)
newday<-c()
for(i in 1:nrow(NewActivity)){
  if(days[i]=='Saturday'|days[i]=='Sunday'){
    newday<-c(newday,'weekend')
  }
  else{
    newday<-c(newday,'weekday')
  }
}
NewActivity<-cbind(NewActivity,newday)
NewActivity$newday<-as.factor(NewActivity$newday)
```

Finally let's plot a time series of average of steps vs interval by week day

```{r,echo=TRUE}
library(lattice)
mediatotal<-aggregate(NewActivity$steps,list(NewActivity$newday,NewActivity$interval),mean,na.rm=TRUE)
colnames(mediatotal)<-c('week','interval','steps')
xyplot(steps~interval|week,data = mediatotal,layout=c(2,1),type='l')
```

